Flowchart of Current Algorithm

### 2008-09-09 docs/flowchart/paper.tex
put a little order into the equations list
starting work on making this accurately reflect the current code

\section{Equations}
\begin{eqnarray}
\frac{dX_k}{dt} &=& \omegadot_k(\rho,X_k,T)\enskip, \label{eq:VODE1C} \\
\frac{dT}{dt} &=&\frac{1}{c_p} \left ( -\sum_k \xi_k  \omegadot_k  + \Hnuc \right )\enskip. \label{eq:tempreactC}
\end{eqnarray}
\begin{equation}
\label{eq:tempenthalpy}
\frac{DT}{Dt} = \frac{1}{\rho c_p} \left[ \left(1 - \rho h_p\right) \frac{D p}{D t}
 - \sum_k \rho \xi_k {\omegadot}_k 
 + \rho \Hnuc + \rho \Hext \right] \enskip , 
\end{equation}
\begin{eqnarray}
\frac{\partial \rhozero}{\partial t} &=& - \nablab \cdotb \left( \rhozero w_0 \er \right)
- \nablab \cdotb \left( \etarho \er \right) \enskip ,
\label{eq:rho0upd_new}
\end{eqnarray}
\begin{equation}
\frac{\partial p_0}{\partial t} = w_0 \rho_0 g  + \etarho g
                                 = - w_0 \frac{\partial p_0}{\partial r} + \etarho g \enskip .
\label{eq:p0upd_new} 
\end{equation}
\begin{equation}
\label{eq:constraint}
\nablab \cdotb \ubold + \alpha \left( \frac{\partial p_0}{\partial t} + \ubold \cdotb \nablab p_0 \right) =
 -\sigma  \sum_k \xi_k \omegadot_k  + 
  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  {\omegadot}_k + \sigma \Hnuc + \sigma \Hext \equiv S \enskip ,
\end{equation}
\begin{eqnarray}
\frac{\partial w_0}{\partial r} &=& \frac{1}{\beta_0} \nablab \cdotb (\beta_0 w_0 \er) 
                                    - w_0 \frac{1}{\beta_0} \frac{\partial \beta_0}{\partial r}  \nonumber \\
                                &=&   \left( \Sbar - \frac{1}{\gammabar p_0}
                                      \frac{\partial p_0}{\partial t} \right)
                                     - w_0 \frac{1}{\beta_0} \frac{\partial \beta_0}{\partial r}  \nonumber \\
                                &=&   \Sbar - \frac{1}{\gammabar p_0}
                                      \left( \frac{\partial p_0}{\partial t}
                                       + w_0 \frac{\partial p_0}{\partial r}  \right) \nonumber \\
                                &=&   \Sbar - \frac{1}{\gammabar p_0} \etarho g \enskip \label{eq:divw0} ,
\end{eqnarray}
\begin{equation}
- \frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} = \frac{\partial w_0}{\partial t} + w_0 \frac{\partial w_0}{\partial r} \enskip, \label{eq:pizero}
\end{equation}
\begin{equation}
\int_{\Omega_H} \ut \cdotb \er \; dA = 0 \label{eq:udoterzero} \enskip ,
\end{equation}
\begin{equation}
\nablab \cdotb (\beta_0 \ut )  = \beta_0 \left(S - \Sbar \right )
 \enskip .  \label{eq:tildeconstraint}
\end{equation}
\begin{equation}
w_0^{n,\star} = \frac{\dt^{n} w_0^{\nmh} + \dt^{n-1} w_0^{\nph,\star}}{\dt^n+\dt^{n-1}}\enskip,\label{eq:w0nstar}
\end{equation}
\begin{equation}
\left(\frac{\partial w_0}{\partial r}\right)^{n,\star} = \frac{1}{\dt^n+\dt^{n-1} } \left [ \dt^{n} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nmh}
+ \dt^{n-1} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nph,\star} \right ] \enskip.\label{eq:dw0drnstar}
\end{equation}
\begin{eqnarray}
\frac{\partial w_0}{\partial t} &=& - w_0 \frac{\partial w_0}{\partial r} - 
\frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} \label{eq:w0upd} \enskip , \\
\frac{\partial\ut}{\partial t} &=& - \ut \cdotb \nablab \ut - w_0 \frac{\partial \ut}{\partial r}
                                 - \left(\ut \cdotb \er\right) \frac{\partial w_0}{\partial r} \er
                                 - \frac{1}{\rho} \nablab\pi
                                 + \frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} \er
                                 - \frac{(\rho-\rhozero)}{\rho} \; g \; \er  \label{eq:utildeupd}  \enskip ,
\end{eqnarray}
\begin{equation}
 L_\beta^\rho \phi =
   D \left ( \beta_0^{\nph} {\bf V} \right) - \beta_0^{\nph} \left(S^{n+1}-\overline{S}^{n+1} \right)\enskip,
\label{eq:nodal_solve}
\end{equation}
\begin{equation}
  S =  -\sigma  \sum_k  \xi_k \omegadot_k  + 
  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  {\omegadot}_k  + \sigma \Hnuc + \sigma \Hext \enskip .
\label{eq:defineS} 
\end{equation}

@@

\item {\bf React State}$(\rho^{\inp},(\rho h)^{\inp},X_k^{\inp},T^{\inp}, (\rho\Hext)^{\inp})\rightarrow$
$(\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp}, T^{\outp}, (\rho \omegadot_k)^{\outp})$ \newline
is the process by which we evolve the species and enthalpy from $X_k^{\inp} \rightarrow X_k^{\outp}$ 
and $(\rho h)^{\inp} \rightarrow (\rho h)^{\outp}$ by solving the following system of equations 
over a time interval of  $\Delta t/2$,
\begin{eqnarray}
\frac{\partial X_k}{\partial t} &=& \omegadot_k\enskip,\\
\frac{\partial (\rho h)}{\partial t} &=& \rho \Hnuc + \rho \Hext\enskip .
\end{eqnarray}
  In particular, to evolve the species, we solve the system:
\begin{eqnarray}
\frac{dX_k}{dt} &=& \omegadot_k(\rho,X_k,T)\enskip, \\
\frac{dT}{dt} &=&\frac{1}{c_p} \left ( -\sum_k \xi_k  \omegadot_k  + \Hnuc \right )\enskip.
\end{eqnarray}
using the stiff ordinary differential equation
 integration methods provided by the {\tt VODE} package (cite VODE).
The absolute error tolerances are set to $10^{-12}$ for the species,
 and a relative tolerance of $10^{-5}$ is used for the temperature.  
The integration yields the new values of the mass fractions, $X_k^{\outp}$.  
Equation (\ref{eq:tempreactC}) is derived from equation (\ref{eq:tempenthalpy}) by assuming that the pressure is constant during the burn state.  
In evolving these equations, we need to evaluate $c_p$ and $\xi_k$.  
In theory, this means evaluating the equation of state for each right-hand side evaluation that {\tt VODE} requires.  
In practice, we freeze $c_p$ and $\xi_k$ at the start of the integration time step and compute them using 
$\rho^{\inp}, X_k^{\inp},$ and $T^{\inp}$ as inputs to the equation of state.  
We note that while temperature is evolved when solving these equations, we do not keep the final temperature, 
nor do we use it to compute the final change in enthalpy.  
Therefore, $T^{\outp} = T^{\inp}$.  Also, note that the density remains unchanged during 
the {\bf React State} step, i.e., $\rho^{\outp} = \rho^{\inp}$.

@@

\item {\bf Advect Base}$(\rho_0^{\inp}, p_0^{\inp}, \beta_0^{\inp}, w_0^{\inp}, \psi^{\inp} ) \rightarrow 
(\rho_0^{\outp}, \rho_0^{\outp,n+\myhalf}, p_0^{\outp})$ 
\newline
is the process by which we update the base state through $\dt$ in time given the
radial velocity $w_0^{\inp}.$ Here we discuss
the algorithm for plane-parallel geometries.  
The base state arrays are all one-dimensional 
in the radial coordinate; we think of the base state quantities as defined
at cell centers, with $w_0$ and $\beta_0$ defined halfway in-between.\\ \\
a)  For the density update, we discretize equation (\ref{eq:rho0upd_new}), neglecting the $\etarho$ term,
to compute the new base state density,
\begin{eqnarray}
\rho_{0,j}^{\outp} &=& \rho_{0,j}^{\inp} - \frac{\dt}{\dr} 
\left [ \left( {{\rho}_0}^{\outp,\nph} {w_0}^{\inp}\right)_{j+\myhalf} -  
         \left( {{\rho}_0}^{\outp,\nph} {w_0}^{\inp}\right)_{j-\myhalf} \right ] \nonumber  \enskip,
\end{eqnarray}
where $j$ refers to the one-dimensional index in the radial direction.
The interface states, ${{\rho}_0}^{\outp,\nph}$, are found via the procedure described in
paper II, Appendix A, and are saved for use later in the overall algorithm.\\ \\
b)  For the pressure update, We discretize equation (\ref{eq:p0upd_new}), 
to compute the new base state pressure,
\begin{eqnarray}
p_{0,j}^\outp &=& p_{0,j}^\inp - \frac{\dt}{2\dr} \left (w_{0,j+\myhalf}^\inp
+ w_{0,j-\myhalf}^\inp \right)
\left (p_{0,j+\myhalf}^{n+\myhalf} -p_{0,j-\myhalf}^{n+\myhalf} \right ) + \dt \; \psi_j^\inp \enskip,
\end{eqnarray}
where the interface states are again found via the procedure described in
paper II, Appendix A.
\\

@@

We now describe the full time advancement algorithm, making frequent
use of the shorthand developed above.  Here, we assume that the
problem is already properly initialized.  We describe the details of
the initialization in \S~ref{Sec:Initialization}.

In paper II we discretized the density and enthalpy evolution equations in 
perturbational form, arguing that this would be less
susceptible to grid effects.  In the present algorithm, the analog would
be to use a perturbational form of the species equations.
Numerical testing has shown that the form of the algorithm presented
here is in fact more robust than the perturbational form 
when the base state mixing terms ($\etarho$) are included.  
In the algorithm described below, density is still 
extrapolated to edges in perturbational form, but 
we treat $X_k$ and $h$ in non-perturbational form.\\

@@

\item Compute the edge states, $(\rho X_k)^{(1),\nph,\star}$ and $(\rho h)^{(1),\nph,\star},$
      for the conservative update of $(\rho X_k)$ and $(\rho h).$  Here we
      predict $\rho'$, $T$, and $X_k$ to the edges, using
      a second-order Taylor expansion in space and time, as described in
      paper II, Appendix A, using $\V = \uadvone+w_0^{\nph,\star} \er$.  
      We explicitly include the reaction terms in the temperature prediction, 
      since we 
      did not update the temperature in {\bf React State}. 
      We use the equation of state and base state density, $\rho_0$,
      to convert these to edge states for $(\rho X_k)$ and $(\rho h)$. 
      We define $\rho^{(1),\nph,\star} = \sum_k (\rho X_k)^{(1),\nph,\star}.$

@@

\item Compute the edge states, $(\rho X_k)^{(1),\nph}$ and $(\rho h)^{(1),\nph},$
      for the conservative update of $(\rho X_k)$ and $(\rho h).$  Here we
      predict $\rho'$, $T$, and $X_k$ to the edges, using
      a second-order Taylor expansion in space and time, as described in
      paper II, Appendix A, using $\V = \uadvtwo+w_0^{\nph} \er$. 
      Again, we explicitly include the reaction terms in the temperature prediction, 
      since we did not update  
      the temperature in {\bf React State}.
      We use the equation of state and base state density, $\rho_0$,
      to convert these to edge states for $(\rho X_k)$ and $(\rho h)$. 
      We define $\rho^{(1),\nph} = \sum_k (\rho X_k)^{(1),\nph}.$

@@

\subsection{Initialization}

>>>>

\section{Equations}

\subsection{Reaction Network}
\begin{eqnarray}
\frac{dX_k}{dt} &=& \omegadot_k(\rho,X_k,T)\enskip, \label{eq:VODE1C} \\
\frac{dT}{dt} &=&\frac{1}{c_p} \left ( -\sum_k \xi_k  \omegadot_k  + \Hnuc \right )\enskip. \label{eq:tempreactC}
\end{eqnarray}

\subsection{Energy}

\begin{equation}
\label{eq:tempenthalpy}
\frac{DT}{Dt} = \frac{1}{\rho c_p} \left[ \left(1 - \rho h_p\right) \frac{D p}{D t}
 - \sum_k \rho \xi_k {\omegadot}_k 
 + \rho \Hnuc + \rho \Hext \right] \enskip , 
\end{equation}

\subsection{Velocity}

\begin{equation}
U = \ut + w_0 \er
\end{equation}

\begin{equation}
\int_{\Omega_H} \ut \cdotb \er \; dA = 0 \label{eq:udoterzero} \enskip ,
\end{equation}

\begin{eqnarray}
\frac{\partial w_0}{\partial t} &=& - w_0 \frac{\partial w_0}{\partial r} - 
\frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} \label{eq:w0upd} \enskip , \\
\frac{\partial\ut}{\partial t} &=& - \ut \cdotb \nablab \ut - w_0 \frac{\partial \ut}{\partial r}
                                 - \left(\ut \cdotb \er\right) \frac{\partial w_0}{\partial r} \er
                                 - \frac{1}{\rho} \nablab\pi
                                 + \frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} \er
                                 - \frac{(\rho-\rhozero)}{\rho} \; g \; \er  \label{eq:utildeupd}  \enskip ,
\end{eqnarray}

\begin{equation}
\label{eq:constraint}
\nablab \cdotb \ubold + \alpha \left( \frac{\partial p_0}{\partial t} + \ubold \cdotb \nablab p_0 \right) =
 -\sigma  \sum_k \xi_k \omegadot_k  + 
  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  {\omegadot}_k + \sigma \Hnuc + \sigma \Hext \equiv S \enskip ,
\end{equation}

\begin{equation}
  S =  -\sigma  \sum_k  \xi_k \omegadot_k  + 
  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  {\omegadot}_k  + \sigma \Hnuc + \sigma \Hext \enskip .
\label{eq:defineS} 
\end{equation}

\begin{equation}
\nablab \cdotb (\beta_0 \ut )  = \beta_0 \left(S - \Sbar \right )
 \enskip .  \label{eq:tildeconstraint}
\end{equation}

\begin{equation}
- \frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} = \frac{\partial w_0}{\partial t} + w_0 \frac{\partial w_0}{\partial r} \enskip, \label{eq:pizero}
\end{equation}

\subsection{Base State}

\begin{eqnarray}
\frac{\partial \rhozero}{\partial t} &=& - \nablab \cdotb \left( \rhozero w_0 \er \right)
- \nablab \cdotb \left( \etarho \er \right) \enskip ,
\label{eq:rho0upd_new}
\end{eqnarray}

\subsubsection{Plane-Parallel}
\begin{equation}
\frac{\partial p_0}{\partial t} = w_0 \rho_0 g  + \etarho g
                                 = - w_0 \frac{\partial p_0}{\partial r} + \etarho g \enskip .
\label{eq:p0upd_new} 
\end{equation}

\begin{equation}
\frac{\partial w_0}{\partial r} =  \Sbar - \frac{1}{\gammabar p_0} \etarho g \enskip \label{eq:divw0} ,
\end{equation}

\subsection{Misc.}

\begin{equation}
w_0^{n,\star} = \frac{\dt^{n} w_0^{\nmh} + \dt^{n-1} w_0^{\nph,\star}}{\dt^n+\dt^{n-1}}\enskip,\label{eq:w0nstar}
\end{equation}
\begin{equation}
\left(\frac{\partial w_0}{\partial r}\right)^{n,\star} = \frac{1}{\dt^n+\dt^{n-1} } \left [ \dt^{n} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nmh}
+ \dt^{n-1} \left(\frac{\partial w_0 }{ \partial r}\right)^{\nph,\star} \right ] \enskip.\label{eq:dw0drnstar}
\end{equation}

\begin{equation}
 L_\beta^\rho \phi =
   D \left ( \beta_0^{\nph} {\bf V} \right) - \beta_0^{\nph} \left(S^{n+1}-\overline{S}^{n+1} \right)\enskip,
\label{eq:nodal_solve}
\end{equation}

@@

\item {\bf React State}$[\rho^{\inp},(\rho h)^{\inp},X_k^{\inp},T^{\inp}, (\rho\Hext)^{\inp}]\rightarrow$
$[\rho^{\outp}, (\rho h)^{\outp}, X_k^{\outp}, T^{\outp}, (\rho \omegadot_k)^{\outp}]$ \newline
is the process by which we evolve the species and enthalpy from $X_k^{\inp} \rightarrow X_k^{\outp}$ 
and $(\rho h)^{\inp} \rightarrow (\rho h)^{\outp}$ by solving the following system of equations 
over a time interval of  $\Delta t/2$,
\begin{eqnarray}
\frac{\partial X_k}{\partial t} &=& \omegadot_k\enskip,\\
\frac{\partial (\rho h)}{\partial t} &=& \rho \Hnuc + \rho \Hext\enskip .
\end{eqnarray}
  In particular, to evolve the species, we solve the system:
\begin{eqnarray}
\frac{dX_k}{dt} &=& \omegadot_k(\rho,X_k,T)\enskip, \\
\frac{dT}{dt} &=&\frac{1}{c_p} \left ( -\sum_k \xi_k  \omegadot_k  + \Hnuc \right )\enskip.
\end{eqnarray}
using the stiff ordinary differential equation
 integration methods provided by the {\tt VODE} package (cite VODE).
The absolute error tolerances are set to $10^{-12}$ for the species,
 and a relative tolerance of $10^{-5}$ is used for the temperature.  
The integration yields the new values of the mass fractions, $X_k^{\outp}$.  
Equation (\ref{eq:tempreactC}) is derived from equation (\ref{eq:tempenthalpy}) by assuming that the pressure is constant during the burn state.  
In evolving these equations, we need to evaluate $c_p$ and $\xi_k$.  
In theory, this means evaluating the equation of state for each right-hand side evaluation that {\tt VODE} requires.  
In practice, we freeze $c_p$ and $\xi_k$ at the start of the integration time step and compute them using 
$\rho^{\inp}, X_k^{\inp},$ and $T^{\inp}$ as inputs to the equation of state.  
We note that while temperature is evolved when solving these equations, we do not keep the final temperature, 
nor do we use it to compute the final change in enthalpy.  
Therefore, $T^{\outp} = T^{\inp}$.  Also, note that the density remains unchanged during 
the {\bf React State} step, i.e., $\rho^{\outp} = \rho^{\inp}$.

@@

\item {\bf Advect Base}$(\rho_0^{\inp}, p_0^{\inp}, \beta_0^{\inp}, w_0^{\inp}, \psi^{\inp} ) \rightarrow 
(\rho_0^{\outp}, \rho_0^{\outp,n+\myhalf}, p_0^{\outp})$ 
\newline
is the process by which we update the base state through $\dt$ in time given the
radial velocity $w_0^{\inp}.$ Here we discuss
the algorithm for plane-parallel geometries.  
The base state arrays are all one-dimensional 
in the radial coordinate; we think of the base state quantities as defined
at cell centers, with $w_0$ and $\beta_0$ defined halfway in-between.\\ \\
a)  For the density update, we discretize equation (\ref{eq:rho0upd_new}), neglecting the $\etarho$ term,
to compute the new base state density,
\begin{eqnarray}
\rho_{0,j}^{\outp} &=& \rho_{0,j}^{\inp} - \frac{\dt}{\dr} 
\left [ \left( {{\rho}_0}^{\outp,\nph} {w_0}^{\inp}\right)_{j+\myhalf} -  
         \left( {{\rho}_0}^{\outp,\nph} {w_0}^{\inp}\right)_{j-\myhalf} \right ] \nonumber  \enskip,
\end{eqnarray}
where $j$ refers to the one-dimensional index in the radial direction.
The interface states, ${{\rho}_0}^{\outp,\nph}$, are found via the procedure described in
paper II, Appendix A, and are saved for use later in the overall algorithm.\\ \\
b)  For the pressure update, We discretize equation (\ref{eq:p0upd_new}), 
to compute the new base state pressure,
\begin{eqnarray}
p_{0,j}^\outp &=& p_{0,j}^\inp - \frac{\dt}{2\dr} \left (w_{0,j+\myhalf}^\inp
+ w_{0,j-\myhalf}^\inp \right)
\left (p_{0,j+\myhalf}^{n+\myhalf} -p_{0,j-\myhalf}^{n+\myhalf} \right ) + \dt \; \psi_j^\inp \enskip,
\end{eqnarray}
where the interface states are again found via the procedure described in
paper II, Appendix A.\\ \\
c)  For the enthalpy update, \MarginPar{We need enthalpy equations now that we are predicting $(\rho h)'$.}
\\

@@

We now describe the full time advancement algorithm, making frequent
use of the shorthand developed above.  Here, we assume that the
problem is already properly initialized.  We describe the details of
the initialization in \S~\ref{Sec:Initialization}.

In paper III we discretized the enthalpy evolution equation in full, non-perturbation form.  
We have discovered that discretizing the enthalpy evolution in non-perturbational form 
leads to better numerical properties.  This is more like paper II.\\

@@

\item Compute the edge states, $(\rho X_k)^{(1),\nph,\star}$ and $(\rho h)^{(1),\nph,\star},$
      for the conservative update of $(\rho X_k)$ and $(\rho h).$  Here we
      predict $\rho'$, $(\rho h)'$, and $X_k$ to the edges, using
      a second-order Taylor expansion in space and time, as described in
      paper II, Appendix A, using $\V = \uadvone+w_0^{\nph,\star} \er$.  
      We do not include the reaction terms in the enthalpy prediction, 
      since we accounted for them already in {\bf React State}. 
      We use the base state density, $\rho_0$, and base state enthalpy, $(\rho h)_0$,
      to convert these to edge states for $(\rho X_k)$ and $(\rho h)$. 
      We define $\rho^{(1),\nph,\star} = \sum_k (\rho X_k)^{(1),\nph,\star}.$

@@

\item Compute the edge states, $(\rho X_k)^{(1),\nph}$ and $(\rho h)^{(1),\nph},$
      for the conservative update of $(\rho X_k)$ and $(\rho h).$  Here we
      predict $\rho'$, $(\rho h)'$, and $X_k$ to the edges, using
      a second-order Taylor expansion in space and time, as described in
      paper II, Appendix A, using $\V = \uadvtwo+w_0^{\nph} \er$. 
      Again, we do not include the reaction terms in the enthalpy prediction 
      since we accounted for them already in {\bf React State}.
      We use the base state density, $\rho_0$, and base state enthalpy, $(\rho h)_0$,
      to convert these to edge states for $(\rho X_k)$ and $(\rho h)$. 
      We define $\rho^{(1),\nph} = \sum_k (\rho X_k)^{(1),\nph}.$

@@

\subsection{Initialization}\label{Sec:Initialization}
